
make a new season's ratings by continuing from last season

## packages

```{r}
library(tidyverse)
library(conflicted)
source("cleaning.R")
```

## begin

get latest ratings

```{r}
with_post <- read_rds("wpost.rds")
with_post
```



grab ukraine

```{r}
all_ratings %>% filter(league=="ukraine") %>% 
  unnest(ratings) -> d
d
```

championship round is 54066, relegation round is 54068


these are final ratings for 54067, so use as initial ratings for 54066-54068.

idea: if current "last" exists as this, get ratings from there, else do old initialization

or: make two rows of data file and combine into list-column

```{r}
d <- tribble(
  ~a, ~b,
  47821, 54067,
  54067, 54066,
  54067, 54068
)
d
```

```{r}
d %>% nest(bb=b) %>% 
  pull(bb)
```

so clean!

```{r}
games %>% filter(comp==54066) %>% 
  clean_names() -> dd
dd
```

```{r}
dd %>% left_join(d, by=c("name"="team"))
```

get new teams (if any) and do like initializing.Rmd

leagues to check

ukraine
austria
armenia
chinese taipei
tajikistan
croatia **
greece

in codes.csv, repeat regular season with each playoff section, then nest this season col when reading in. Deal with consequences, eg first element of list, %in% vs ==, etc.