---
title: "Estimating"
output: html_notebook
---

## startup

```{r}
library(tidyverse)
source("functions.R")
source("cleaning.R")
source("estimation.R")
```
```{r}
warnings()
```




keep track of league codes in codes.csv (libre office)

```{r}
codes <- read_csv("codes.csv")
codes
```





idea: do a combined log-like for a collection of values of k (say seq(0, 100, 5)). That is, do it for each league and sum by value.

```{r}
ks <- seq(5, 20, 5)
codes %>% sample_frac(1) %>% 
  crossing(k=ks) -> d
pb <- progress_estimated(nrow(d))
d %>% 
  mutate(ll=pmap_dbl(list(k, last, this), ~est_k(..1, ..2, ..3))) -> dd
dd
```

```{r}
dd %>% 
  mutate(k=sprintf("k%02d", k)) %>% 
  pivot_wider(names_from=k, values_from=ll) %>% 
  summarize_at(vars(starts_with("k")), ~sum(.))
```

this suggests 10 points is better than 20 (that I have seen elsewhere). There is only one parameter k, so anything within 1.92 of the max likelihood is within a 95% CI, right?

all right, this says k should be 10. Can I construct a df of teams and ratings?

## Teams and ratings

```{r}
codes <- read_csv("codes.csv")
codes %>% filter(is.na(skip)) %>% select(-skip) %>% sample_frac(1) -> codes
codes
```

select a few

```{r}
codes %>% sample_n(3) -> c3
c3
```


```{r}
c3 %>% mutate(preseason=map2(last, this, ~preseason_ratings(.x, .y, games))) -> c3
c3
```

this season

```{r}
c3 %>% mutate(post=map2(this, preseason, ~this_ratings(.x, .y, 10, games))) -> c3
c3 %>% mutate(rat=map(post, "r")) %>% unnest(rat) %>% arrange(league, desc(rating)) %>% 
  select(league, team, rating, h) %>% View()
```

```{r}
c3 %>% mutate(the_games=map(post, "d")) %>% 
  unnest(the_games) %>% 
  arrange(desc(time_stamp)) %>% 
  select(-last, -this, -preseason, -post) %>% View()
```


this is good. Get rating table sorted, and most recent games with deltas.



## jottings

## let's do Scotland this season

first, get clean games from last season and estimate initials

```{r}
scotland <- read_rds("scotland.rds")
scotland
```

```{r}
estimate_initial(scotland) -> scotland_initial
scotland_initial
```

read cleaned this-season games

```{r}
comp=53513
(get_clean_games(comp, games) -> scotland_current)
```

make a list of teams in this one

```{r}
with(scotland_current, unique(c(t1, t2))) %>% enframe(name=NULL, value="team") -> teams_this
teams_this
```

find missing/extra ones

```{r}
teams_this %>% anti_join(scotland_initial, by=c("team"="name")) %>% pull(team) -> up # came up
scotland_initial %>% anti_join(teams_this, by=c("name"="team")) %>% pull(name) -> down # went down
tibble(up, down) -> changes
changes
```

can I automate the replacement of team name?

```{r}
scotland_initial %>% left_join(changes, by=c("name"="down")) %>% 
  mutate(name=ifelse(!is.na(up), up, name)) %>% 
  select(-up) -> scotland_initial
scotland_initial
```

going back to basics:

```{r}
ratings <- scotland_initial$rat
names(ratings) <- scotland_initial$name
ratings # named vector for looking up team names
h <- scotland_initial$h[1]
h
```

a big loop through this year's games

```{r}
scotland_current
```

```{r}
k <- 40
scotland_current %>% mutate(p=NA, delta=NA) -> scotland_current
scotland_current
for (i in 1:nrow(scotland_current)) {
  t1name <- scotland_current$t1[i]
  t2name <- scotland_current$t2[i]
  result <- scotland_current$r[i]
  rat1 <- ratings[t1name]
  rat2 <- ratings[t2name]
  diff <- rat1-rat2+h
  p <- 1/(1+10^(-diff/400))
  delta <- k*(result-p)
  ratings[t1name] <- ratings[t1name] + delta
  ratings[t2name] <- ratings[t2name] - delta
  h <- h + delta/10
  scotland_current$p[i] <- p
  scotland_current$delta[i] <- delta
}
scotland_current
```

```{r}
new_ratings(40, scotland_current, scotland_initial)
```

optimizing k

need a function that returns the negative log likelihood (using that column p)

```{r}

log_lik(0, scotland_current, scotland_initial)
```

try some values

```{r}
v <- tibble(k=seq(-10, 10, 1))
v %>% mutate(loglik=map_dbl(k, ~log_lik(., scotland_current, scotland_initial)))
```

the optimal k appears to be *negative*!

try to optimize anyway

```{r}
ans <- optimize(log_lik, c(0, 60), d=scotland_current, r=scotland_initial)
ans
```

now, how about a function that starts from two comp numbers, optimizes for k, and then estimates new ratings?

