
---
title: "elo"
output: html_notebook
---

I can add a readme

## packages

```{r}
library(tidyverse)
source("functions.R")
source("cleaning.R")
```


## initial ratings as "static elo"

[ref](http://opisthokonta.net/?p=1412)

idea: estimate elo ratings by maximum likelihood (or, alternatively, using Stan maybe, "optimizing")

benefit to this idea is that it's all base + tidyverse

```{r}
games <- readRDS("~/teaching/scoresway/rds/games.rds")
```


```{r}
comp=53499
ei_comp(comp, games) %>% arrange(desc(rat))
```


## jottings

that is, estimate the (unchanging) ELO ratings that max the likelihood

try a test one

```{r}
d <- tribble(
  ~t1, ~t2, ~r,
  "a", "b", 1,
  "b", "a", 0.5,
  "a", "b", 1,
  "b", "a", 0.5
)
d
```

get distinct teams and make lookup table

```{r}
v <- with(d, fct_c(factor(t1), factor(t2)))
enframe(levels(v), name="id", value="name") -> team_lookup
team_lookup
```

augment d with lookup numbers

```{r}
d %>% left_join(team_lookup, by=c("t1"="name")) %>% 
  left_join(team_lookup, by=c("t2"="name")) -> dd
dd
```

going to use `optim` shortly

function with input ratings (as vector?) and work out likelihood

```{r}
rat <- c(1700, 1500, 10)
nt <- nrow(team_lookup)
dd %>% mutate(d=rat[id.x]-rat[id.y]+rat[nt+1]) %>% 
  mutate(p=elo_prob(d)) %>% 
  mutate(loglik=r*log(p)+(1-r)*log(1-p)) %>% 
  summarize(ll=sum(-2*loglik)) %>% pull(ll)
```

now arrange into function. Input rat (as vector), data frame of games, data frame of team name lookup

```{r}
make_team_lookup <- function(games) {
  v <- with(games, fct_c(factor(t1), factor(t2)))
  enframe(levels(v), name="id", value="name")
}

mloglik <- function(rat, games, team_lookup) { # minus twice loglik
  # team lookups
  d %>% left_join(team_lookup, by=c("t1"="name")) %>% 
    left_join(team_lookup, by=c("t2"="name")) -> dd
  nt <- nrow(team_lookup)
  dd %>% mutate(d=rat[id.x]-rat[id.y]+rat[nt+1]) %>% 
    mutate(p=elo_prob(d)) %>% 
    mutate(loglik=r*log(p)+(1-r)*log(1-p)) %>% 
    summarize(ll=sum(-2*loglik)) %>% pull(ll)
}
```

test

```{r}
rat <- c(1600, 1500, 10)
mloglik(rat, d, team_lookup)
```

regularization

```{r}
r <- c(12, 11, 10)
v <- 7

```


```{r}
m2 <- function(rat0, games, team_lookup) {
  # rat is n-1 independent ratings with the last one being the home adv
  n <- length(rat0)
  h <- rat0[n]
  target_total <- 1500*n
  actual_total <- sum(rat0[1:(n-1)])
  rat0[n] <- target_total-actual_total
  rat <- c(rat0, h)
  mloglik(rat, games, team_lookup)
}
```

function to control this. I am estimating elo ratings from afresh:


```{r}
estimate_initial <- function(games, team_lookup) {
  n <- nrow(team_lookup)
  rat0 <- c(rep(1500, n-1), 0)
  ans <- optim(rat0, fn=m2, games=d, team_lookup=team_lookup)
  rat0 <- ans$par
  # reconstitute
  h <- rat0[n]
  target_total <- 1500*n
  actual_total <- sum(rat0[1:(n-1)])
  rat0[n] <- target_total - actual_total
  list(rat=rat0, h=h)
}
```

try this

```{r}
estimate_initial(games=d, team_lookup=team_lookup)
```

set up some games to estimate from

```{r}
d <- tribble(
  ~t1, ~t2, ~r,
  "a", "b", 1,
  "b", "c", 1,
  "c", "a", 0,
  "b", "a", 0.5,
  "c", "b", 0.5, 
  "a", "c", 1
)
d
```


estimate

```{r}
estimate_initial(games=d)
```

try it with something real

```{r}
games <- readRDS("~/teaching/scoresway/rds/games.rds")
games %>% filter(comp==47730) %>% 
  select(t1=t1_name, t2=t2_name, score) %>% 
  filter(str_detect(score, " - ")) %>% 
  separate(score, into=c("s1", "s2"), convert = T) %>% 
  mutate(r=case_when(
    s1>s2 ~ 1,
    s1<s2 ~ 0,
    TRUE  ~ 0.5
  )) -> d
d
```

```{r}
estimate_initial(d) %>% arrange(desc(rat))
```

or clean the team names first

```{r}
games <- readRDS("~/teaching/scoresway/rds/games.rds")
games %>% filter(comp==47730) -> scotland
cleaned_names <- clean_names(scotland)
lookup_cleaned_names(scotland, cleaned_names) %>% 
  filter(str_detect(score, " - ")) %>% 
  arrange(time_stamp) %>% 
  separate(score, into=c("s1", "s2"), convert = T) %>% 
  mutate(r=case_when(
    s1>s2 ~ 1,
    s1<s2 ~ 0,
    TRUE  ~ 0.5
  )) -> d
d
estimate_initial(d)

```




